\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}

\newcommand{\PEGSPending}[1]{\textbf{[PENDING: #1]}}

\title{Detectability and Warning Value of Prompt Elasto-Gravity Signals from Manila Trench Megathrust Earthquakes}
\author{\PEGSPending{author list, station operators, and warning partners}}
\date{Methods-first draft; not for submission}

\begin{document}
\maketitle

\begin{abstract}
We evaluate when prompt elasto-gravity signals (PEGS) from Manila Trench
megathrust scenarios can be detected and used for reliable earthquake
characterization under realistic South China and surrounding station geometry,
non-Gaussian continuous noise, high-microseism and typhoon conditions, station
outages, and explicit operational false-alarm constraints. The study compares
single-station and coherent physical baselines with a conditional graph model,
and quantifies decision-time gain relative to P-wave, W-phase, and GNSS
characterization using scenario- and location-specific tsunami arrivals.
\PEGSPending{Insert registered simulator, station, noise, baseline, and holdout
results.}
\end{abstract}

\section{Introduction}
PEGS theory, empirical-noise graph models, and a simulated real-time tsunami-
warning implementation already exist. The question addressed here is narrower:
whether South China and surrounding networks can provide reliable Manila Trench
decisions at a fixed operational false-alarm rate, and what station/noise
requirements define the feasible boundary.

\section{Scenario and station data}
Each source record binds segment, Mw, strike/dip/rake, depth, rupture dimensions,
slip, rise time, rupture velocity, and publication. Every tsunami arrival is
nested under that source and has a location and arrival definition. Qualitative
statements such as ``more than about three hours'' are not converted into exact
universal arrival times.

The open Zhao and Niu scenario descriptor~\cite{Zhao2025} separates southern
(13.0--14.5$^\circ$N) and northern (14.5--22.0$^\circ$N) Manila source families,
with 1000-year recurrence-family magnitudes $M_w$ 8.1 and 9.1. Its 1,800
0.1-degree Slab2.0 subfaults provide location, depth, strike and dip, and the
hazard calculation assumes 90-degree rake. The released data are target-point
height/period statistics, not per-scenario PEGS sources: rise time, rupture
velocity, complete rupture geometry and location-specific arrivals remain
missing. These source families are not promoted to simulator inputs.

Liu et al.~\cite{Liu2023} provide a complementary PMEL/SIFT-based set of
19 finite-fault families: eleven $M_w$ 7.5, five $M_w$ 8.1 and three $M_w$ 8.5
locations, with reported rupture dimensions, mean slip, rake and individual
static geometry. Their arrival ranges remain tied to magnitude family and site
(Nanao, Shanwei, Macao, Boao or Sanya). Rise time and rupture velocity are not
reported, so these records also remain family evidence rather than PEGS-ready
dynamic scenarios.

Station candidates require a contemporaneous BH/LH three-component epoch, full
response, waveform availability, licence, latency class, and noise quality.
Channel metadata or scalar sensitivity alone is insufficient. Noise windows are
frozen by season, day/night, calm sea, typhoon/microseism, urban/remote setting,
and outage state before model fitting.

The archive/response inventory is retrieved through EarthScope FDSN
services~\cite{EarthScopeFDSN}. Network-specific citations and acknowledgements
remain mandatory; archive availability is never relabelled as real-time latency.

\section{Methods}
\subsection{PEGS simulation benchmark}
The selected simulator must reproduce a published event at specified stations
and components, with predeclared waveform correlation, amplitude, and onset
tolerances, before Manila scenarios are generated. Earth model, reference frame,
source-time function, units, sampling, and code commit are recorded.

\subsection{Splits and domain shift}
Training, validation, and test sets separate rupture segments and noise dates.
Domain-out sources and stations are retained. Simultaneous station noise windows
preserve spatial correlation. Synthetic white noise is an ablation, not the
primary evaluation.

\subsection{Physical baselines}
Baselines proceed from single-station energy to explicit phase/time-aligned
multi-station coherent stacking, template/likelihood detection, and rapid source
parameter inversion. Station weights are explicit and L2-normalized. Fixed 0\%,
20\%, and 40\% outage masks retain exact station identities.

\subsection{False alarms and detection probability}
Noise-only scores calibrate a threshold in false alarms per 30-day month. If the
record is too short to resolve one exceedance at the target rate, the threshold
is placed above the observed maximum and the resolution limit is reported.
Held-out detection probability is then evaluated versus time. Earliest reliable
detection may require multiple consecutive qualifying time points.

\subsection{Reliable magnitude and warning value}
At every time we report magnitude MAE and bias, sensitivity for the predeclared
high-risk class (nominally $M_w\geq8.2$), lower-risk specificity, and prediction-
interval coverage. Reliable magnitude time is the first sustained point meeting
all criteria.

For location $x$,
\begin{align}
 \Delta t_{\mathrm{characterization}} &=
 t_{\mathrm{conventional,reliable\ Mw}}-t_{\mathrm{PEGS,reliable\ Mw}},\\
 T_{\mathrm{lead}}(x) &= t_{\mathrm{tsunami\ arrival}}(x)
 -t_{\mathrm{PEGS,reliable\ Mw}}.
\end{align}
Negative characterization gain is retained as a negative result.

\subsection{Network optimization and conditional GNN}
Existing, idealized, and incrementally augmented networks are compared by
detection time/probability, false alarms, magnitude error, robustness, and
station count. A graph model is trained only if held-out real-noise physical
baselines demonstrate stable information. It must outperform those baselines
under identical splits and thresholds, not merely fit synthetic waveforms.

\section{Results}
A registered open-archive audit exact-matches five three-component LH station
epochs to complete response structure: HK.HKPS, IC.QIZ, IU.DAV, MY.KKM and
TW.KMNB. Their common audited archive interval begins in 2020. The frozen outage
design contains one five-station, five four-station and ten three-station
combinations for nominal 0\%, 20\% and 40\% branches.

An initial real-noise pilot requested three simultaneous six-hour windows.
Thirteen of fifteen waveform requests and twelve response-removal records pass;
the failures are retained as two KKM 404 responses and one KMNB gap/overlap
case. A failure-triggered replacement adds one complete five-station window.
Across the two complete unclassified windows, maximum absolute off-diagonal
vertical correlation is 0.039 and 0.311, respectively. These are descriptive
archive/response-QC results. No window is yet authorized as quiet calibration
data, and no PEGS signal was evaluated. \PEGSPending{Insert detection, magnitude
and warning results only after the simulator benchmark and environment-labelled
noise gates pass.}

\section{Discussion}
A high-value negative result is an explicit station-count, placement, noise, and
magnitude boundary. Preceding tsunami arrival alone does not establish decision
value; the system must characterize hazard reliably and rarely enough to be
operationally usable.

\section{Data and code availability}
Station metadata, waveform licences, exact noise dates, simulator/code versions,
scenario provenance, splits, thresholds, compact metrics and checksums will be
recorded. Restricted waveforms and large synthetic corpora remain outside Git.

\section{Limitations}
Archive availability is not real-time latency. The present acceleration spectra
use a diagnostic 0.005--0.05~Hz band for response QC, not a frozen PEGS detection
passband. Candidate calibration/held-out labels still require earthquake,
weather, microseism and transient audits. The simulator benchmark, physical
detectability boundary, network performance, warning value and every graph-model
stage remain pending. The extracted source families also lack per-scenario rise
time, rupture velocity, complete geometry and location-specific arrivals.

\bibliographystyle{plain}
\bibliography{references}
\end{document}
