\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}

\newcommand{\TyphoonPending}[1]{\textbf{[PENDING: #1]}}

\title{Event-Resolved Gravimetric Detection of Typhoon-Driven Ocean Mass Redistribution}
\author{\TyphoonPending{author list and data-provider collaboration}}
\date{Methods-first draft; not for submission}

\begin{document}
\maketitle

\begin{abstract}
We test whether named typhoons produce event-resolved ocean mass-redistribution
signals in superconducting-gravimeter records after explicit separation of
atmosphere, hydrology, direct ocean attraction, elastic loading, and instrument
effects. The design combines gravimetry with pressure, best tracks, gridded sea
level, tide gauges and GNSS, and evaluates a completely held-out typhoon together
with strong-weather and quiet controls. \TyphoonPending{Insert data-gated
multi-event results and uncertainty after station authorization.}
\end{abstract}

\section{Introduction}
Previous work has modeled non-tidal ocean loading in the South China Sea and
observed storm-surge loading at Helgoland. The increment tested here is not
general NTOL correction, but named-typhoon attribution with a closed component
budget, independent evidence, and cross-event generalization.

\section{Data and preregistered event design}
The raw-data gate requires either one station covering at least three typhoons or
two stations covering at least two typhoons each. Required inputs include
gravity, colocated pressure, calibration and maintenance logs, timing and jump
records, sea-level fields, storm tracks, precipitation, and hydrology.
Evaluation additionally requires one
held-out typhoon, one non-typhoon strong-weather control, three quiet windows,
and no overlapping same-station windows across splits.

IBTrACS agency wind and pressure fields are retained separately. ERA5 surface
pressure supplies atmospheric loading; mean-sea-level pressure and winds are
event diagnostics. Significant wave height is not converted into additional
water mass. The frozen CMEMS MOWL product separates dynamic sea level, inverse
barometer, tide, steric and mass-volume contributions~\cite{CMEMSGLO024}. We
select its \texttt{sea\_surface\_height} component, which excludes tides and
surface-pressure forcing, and add the ERA5-driven inverse-barometer response
exactly once.

\section{Methods}
\subsection{Transparent gravimeter correction}
For observed gravity $g_{\mathrm{obs}}$, the residual is
\begin{equation}
 g_{\mathrm{res}} = g_{\mathrm{obs}}-
 \sum_k g_k,
\end{equation}
where every removed component has a unique physical-effect identifier, source,
SI series, and pre-applied flag. Two components cannot own the same effect. Each
stage stores its input, removed series, output, and peak contribution.

Missing samples and cadence violations split records without interpolation.
Candidate discontinuities are flagged at original indices but are not
automatically classified or corrected.

\subsection{Independent ocean forward model}
Sea-level anomaly $\eta$ is converted to surface density
$\Delta\sigma=\rho_w\eta$. Direct attraction is integrated on spherical and
WGS~84 grids with explicit masks and partial coastal-cell fractions. The modeled
ocean signal is
\begin{equation}
 g_{\mathrm{ocean}}^{\mathrm{model}}=
 g_{\mathrm{direct}}+g_{\mathrm{deformation}}+g_{\mathrm{internal}},
\end{equation}
with every term retained separately. Elastic terms are disabled until the
Green-function Earth model, reference frame, normalization, and component
semantics pass a published benchmark.

\subsection{Double-count audit}
A CMEMS field containing inverse-barometer response cannot be combined with a
second ERA5-driven inverse-barometer correction. Likewise, a pre-corrected NTOL
residual cannot receive NTOL again. Unknown product physics blocks assembly.

\subsection{Evaluation}
Primary metrics are correlation, RMSE, peak amplitude and timing error,
explained variance, spectral coherence, event SNR, and held-out performance.
Quiet and control windows estimate false positives. Sensitivity tests cover the
model, filters, hydrology, Green functions, stations, and events; failed events
are retained.

\section{Results}
A registered non-restricted readiness audit connects 17 completed processing
and evaluation work units. The open IBTrACS inventory contains 48 broad regional
track candidates and 11 priority station-coverage inquiries. These are not
gravity events and do not select a pilot: the real-event gate remains
\texttt{pending\_no\_event\_windows}. A second registered audit closes all six
environmental-effect ownership entries. In particular, CMEMS dynamic sea level
is the ocean direct/elastic input while the CMEMS inverse-barometer component is
excluded in favour of the ERA5-derived branch. This removes a double-counting
risk but supplies no gravity observation. Consequently the claim-safe branch is
\texttt{pending\_evidence}. \TyphoonPending{Insert only registered real-SG
event results after the station-data gate passes.}

\section{Discussion}
Possible outcomes include successful event attribution, evidence that ocean
products miss rapid coastal redistribution, or an upper constraint from
non-detection. A single successful event cannot establish generality.

\section{Data and code availability}
Restricted SG and station logs will remain outside Git. Versioned manifests will
record owner, terms, station, channels, time range, product version, retrieval
time, checksum, exclusions, and redistribution limits. Processing code and
compact nonrestricted metrics will be released.

\section{Limitations}
SG event windows, calibration and station logs remain authorization-dependent.
CMEMS pressure-response semantics and correction ownership are now closed at
the metadata level, but event-specific CMEMS/ERA5 retrieval and anomaly
construction have not occurred. Consequently no pilot event, attribution
coefficient, held-out improvement, event SNR or false-positive rate is reported.
The current method audit cannot substitute for those observations.

\bibliographystyle{plain}
\bibliography{references}
\end{document}
